---
title: "Deprivation and energy consumption"
author: "Stefano De Sabbata"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggrepel)
```


## Data

This document uses data that contain public sector information licensed under the [Open Government Licence v3.0](https://www.nationalarchives.gov.uk/doc/open-government-licence/version/3/) from the [Department for Business, Energy & Industrial Strategy](https://www.gov.uk/government/organisations/department-for-business-energy-and-industrial-strategy), the [Ministry of Housing, Communities & Local Government](https://www.gov.uk/government/organisations/ministry-of-housing-communities-and-local-government) and the [Office for National Statistics](https://www.ons.gov.uk/) (see also [ONS geography licences](https://www.ons.gov.uk/methodology/geography/licences)).



### Load data

```{r, message=FALSE, warning=FALSE}
imd_2019 <- read_csv("../data/indices-of-multiple-deprivation-2019.csv")
elec_2010_2020 <- read_csv("../data/LSOA_domestic_elec_2010-20.csv")
gas_2010_2020 <- read_csv("../data/LSOA_domestic_gas_2010-20.csv")
```

### Merge data

```{r, message=FALSE, warning=FALSE}
energy_deprivation <-
  imd_2019 %>% 
  filter(str_starts(`Local Authority District code (2019)`, "E")) %>% 
  select(-`LSOA name (2011)`) %>% 
  rename(`LSOA code` = `LSOA code (2011)`) %>% 
  left_join(
    elec_2010_2020 %>% 
      left_join(gas_2010_2020)
  )
```

## Visualisation

### Median consumption

```{r}
plot_england <-
  energy_deprivation %>% 
  filter(
    Year == 2020
  ) %>% 
  group_by(`Local Authority District name (2019)`) %>% 
  summarise(
    `Median gas consumption (kWh per meter)` = median(`Median gas consumption (kWh per meter)`),
    `Median electricity consumption (kWh per meter)` = median(`Median electricity consumption (kWh per meter)`)
  ) %>% 
  mutate(
    `Selected LAD` = if_else(
      `Local Authority District name (2019)` == "Leicester",
      `Local Authority District name (2019)`, "Other LAD"
      ),
    plot_label_leicester = if_else(
      `Local Authority District name (2019)` == "Leicester",
      "Leicester", as.character(NA)
    )
  ) %>% 
  ggplot(aes(
    x = `Median gas consumption (kWh per meter)`,
    y = `Median electricity consumption (kWh per meter)`,
    fill = `Selected LAD`
  )) +
  geom_text_repel(
    aes(label = plot_label_leicester),
    colour = "#003090",
    xlim = c(14000, NA),
    ylim = c(NA, 2600),
    force = 1,
    segment.curvature = -0.1
  ) + 
  geom_point(shape = 21, colour = "#666666", alpha = 0.7) +
  scale_fill_manual(values = c("#003090", "#cccccc")) +
  guides(
   fill = guide_legend(
     title = ""
   )
  ) +
  xlim(0, 25000) +
  ylim(1500, 4500) +
  ggtitle("Overview in England") +
  theme_bw()
```

```{r}
plot_leicester <-
  energy_deprivation %>% 
  filter(
    `Local Authority District name (2019)` == "Leicester" &
    Year == 2020
  ) %>% 
  ggplot(aes(
    x = `Median gas consumption (kWh per meter)`,
    y = `Median electricity consumption (kWh per meter)`,
    fill = as_factor(`Index of Multiple Deprivation (IMD) Decile (where 1 is most deprived 10% of LSOAs)`)
  )) +
  geom_point(shape = 21, colour = "#666666", alpha = 0.7) +
  scale_fill_viridis_d(option = "magma") +
  guides(
    fill = guide_legend(
      title = "IMD 2019"
    )
  ) +
  xlim(0, 25000) +
  ylim(1500, 4500) +
  ggtitle("Detail in Leicester") +
  theme_bw()
```

```{r}
plot_england
```

```{r}
plot_leicester
```

### Mean consumption

```{r}
plot_england <-
  energy_deprivation %>% 
  filter(
    Year == 2020
  ) %>% 
  group_by(`Local Authority District name (2019)`) %>% 
  summarise(
    `Mean gas consumption (kWh per meter)` = mean(`Mean gas consumption (kWh per meter)`),
    `Mean electricity consumption (kWh per meter)` = mean(`Mean electricity consumption (kWh per meter)`)
  ) %>% 
  mutate(
    `Selected LAD` = if_else(
      `Local Authority District name (2019)` == "Leicester",
      `Local Authority District name (2019)`, "Other LAD"
      ),
    plot_label_leicester = if_else(
      `Local Authority District name (2019)` == "Leicester",
      "Leicester", as.character(NA)
    )
  ) %>% 
  ggplot(aes(
    x = `Mean gas consumption (kWh per meter)`,
    y = `Mean electricity consumption (kWh per meter)`,
    fill = `Selected LAD`
  )) +
  geom_text_repel(
    aes(label = plot_label_leicester),
    colour = "#003090",
    xlim = c(16000, NA),
    ylim = c(NA, 3300),
    force = 1,
    segment.curvature = -0.1
  ) + 
  geom_point(shape = 21, colour = "#666666", alpha = 0.7) +
  scale_fill_manual(values = c("#003090", "#cccccc")) +
  guides(
   fill = guide_legend(
     title = ""
   )
  ) +
  xlim(0, 27500) +
  ylim(2000, 6000) +
  ggtitle("Overview in England") +
  theme_bw()
```

```{r}
plot_leicester <-
  energy_deprivation %>% 
  filter(
    `Local Authority District name (2019)` == "Leicester" &
    Year == 2020
  ) %>% 
  ggplot(aes(
    x = `Mean gas consumption (kWh per meter)`,
    y = `Mean electricity consumption (kWh per meter)`,
    fill = as_factor(`Index of Multiple Deprivation (IMD) Decile (where 1 is most deprived 10% of LSOAs)`)
  )) +
  geom_point(shape = 21, colour = "#666666", alpha = 0.7) +
  scale_fill_viridis_d(option = "magma") +
  guides(
    fill = guide_legend(
      title = "IMD 2019"
    )
  ) +
  xlim(0, 27500) +
  ylim(2000, 6000) +
  ggtitle("Detail in Leicester") +
  theme_bw()
```

```{r}
plot_england
```

```{r}
plot_leicester
```


## Session info

```{r}
sessionInfo()
```

